---
title: "Assignment/Explore Query Planning and Indexing"
author: Balakrishna,D
date: Summer Full 2023
output: pdf_document
editor_options: 
  chunk_output_type: console
---

## Establishing DB Connection

```{r Connection Establishment}
library(DBI)
library(RSQLite)

# Connect to the SQLite database
conn <- dbConnect(RSQLite::SQLite(), "sakila.db")
```

## Question 1

```{r Question 1}
# Drop user-defined Indexes
dbExecute(conn, "DROP INDEX TitleIndex;")

# Get the number of films per language
query <- "SELECT l.name, COUNT(*) AS num_films FROM film f LEFT JOIN language l ON
f.language_id=l.language_id GROUP BY f.language_id;"
results <- dbGetQuery(conn, query)

# Print the results
print(results)
```

## Question 2

```{r Question 2}

# Get the query plan for the query
query <- "EXPLAIN QUERY PLAN SELECT l.name, COUNT(*) AS num_films 
          FROM film f LEFT JOIN language l ON f.language_id=l.language_id 
          GROUP BY f.language_id;"

query_plan <- dbGetQuery(conn, query)

# Print the query plan
print(query_plan)
```

## Question 3

The following SQL query will return the title, category name and length of the film with the title "ZORRO ARK":

```{r Question 3}

# Get the title, category name and length of the film with the title "ZORRO ARK"
query3 <- "SELECT f.title AS title, c.name AS category, f.length AS length 
          FROM film f JOIN film_category fc ON f.film_id=fc.film_id JOIN 
          category c ON fc.category_id=c.category_id WHERE title = 'ZORRO ARK';"
results <- dbGetQuery(conn, query3)

# Get the time taken for execution
query_3_time <- system.time(dbGetQuery(conn, query3))

# Print the results
print(results)


```

## Question 4

The query plan for the query in step 3 is as follows:

```{r Question 4}

# Get the query plan for the query
query <- "EXPLAIN QUERY PLAN SELECT f.title AS title, c.name AS category, 
          f.length AS length FROM film f JOIN film_category fc 
          ON f.film_id=fc.film_id JOIN category c ON 
          fc.category_id=c.category_id WHERE title = 'ZORRO ARK';"
query_plan <- dbGetQuery(conn, query) 

# Print the query plan
print(query_plan)
```

## Question 5

The following SQL statement will create a user-defined index called "TitleIndex" on the column TITLE in the table FILM:

```{r Question 5}

# Create a user-defined index called "TitleIndex" on the column `TITLE` 
# in the table `FILM`

query <- "CREATE INDEX TitleIndex ON film (title);"
dbExecute(conn, query)
```

## Question 6

```{r Question 6}
# Re-run the query from step 3 now that you have an index

query6 <- "SELECT f.title AS title, c.name AS category, f.length AS length 
            FROM film f JOIN film_category fc ON f.film_id=fc.film_id 
            JOIN category c ON fc.category_id=c.category_id 
            WHERE title = 'ZORRO ARK';"
results <- dbGetQuery(conn, query6)

# Get the time taken for execution
query_6_time <- system.time(dbGetQuery(conn, query6))

# Print the results
print(results)

query <- "EXPLAIN QUERY PLAN SELECT f.title AS title, c.name AS category, 
          f.length AS length FROM film f JOIN film_category fc 
          ON f.film_id=fc.film_id JOIN category c 
          ON fc.category_id=c.category_id 
          WHERE title = 'ZORRO ARK';"
query_plan <- dbGetQuery(conn, query) 

# Print the query plan
print(query_plan)
```

## Question 7

The query plans for both query(4) and query(6) are slightly different. The difference is notably scene in the detail column of the plans. In query(4), the table was searched using the primary key, whereas after creating the index, we can see in the query plan of query(6) that the table is searched using the index we defined.

## Question 8

Execution times for both query(3) and query(6):

```{r Question 8}
# Print the execution times
print("Query 3: ")
print(query_3_time)
print("Query 6: ")
print(query_6_time)
```

Here we can clearly see that indexing improves execution times for the same results which are obtained from the same queries.The elapsed time has clearly decreased.

## Question 9

```{r Question 9}
# Get the title, language and length of all films with the word "GOLD" with any 
# capitalization in its name

query <- "SELECT f.title AS title, l.name AS language, f.length AS length 
          FROM film f LEFT JOIN language l ON f.language_id=l.language_id 
          WHERE title LIKE '%GOLD%';"
results <- dbGetQuery(conn, query)

# Print the results
print(results)
```

## Question 10

Here we get the query plan for query(9):

```{r Question 10}
# Get the query plan for the query
query <- "EXPLAIN QUERY PLAN SELECT f.title AS title, l.name AS language, 
          f.length AS length FROM film f LEFT JOIN language l 
          ON f.language_id=l.language_id WHERE title LIKE '%GOLD%';"
query_plan <- dbGetQuery(conn, query)

# Print the query plan
print(query_plan)
```
